\newcommand{\popasm}{\emph{PopAsm}}
\newcommand{\nasm}{\emph{NASM}}
\newcommand{\tasm}{\emph{TASM}}

\documentclass[a4paper,draft,12pt]{book}
\title{\popasm\ --- The Popular Assembler\\Developer's Manual}
\author{Helcio B. de Mello}
\date{\today}

\begin{document}

\begin{titlepage}
\maketitle
\end{titlepage}

\pagenumbering{roman}

\bf{NOTE:} This is but a draft of the \popasm\ Developer's Manual.
The contents of this document may change as needed, until the release
of the first official full version. Comments on English mistakes
or contents are welcome, given the author's native language is not
English. Please see appendix \ref{CONTACT} for contacting info.

\chapter*{About this document}
This document is part of \popasm, the Popular Assembler Project.
It has been totally written in \LaTeX\ by
\popasm\ author himself.

At the time of this writing, \popasm\ had two manuals:

\begin{itemize}
\item{User's Manual} --- Aimed to those who wish to \emph{use}
\popasm, regardless the way it was developed and its internal
implementation. This document also concerns about instalation
procedures.
\item{Programmer's Guide} --- This document. Documents \popasm\
source code in detail. It is recommended to those who want to
read, understand and/or modify \popasm\ sources.
\end{itemize}

Please check which manual is the one intended for your needs.

\include{author}
\include{acks}

\tableofcontents
\newpage
\listoftables
\newpage
\listoffigures
\newpage

\pagenumbering{arabic}
\part{Getting Started}
This initial part starts with an introductory chapter which gives an
overall introduction of how \popasm\ code is organized, so the reader
can identify which of the chapters ahead will be of his interest.

The remaining chapters deal with the main classes used by \popasm\ 
to perform all of its functions. They are intended to those who
wish to understand how things work at the highest level of abstraction,
ignoring implementation details of the base classes (eg.: how two
numbers are added is not a concern in this part).

\chapter{Introduction}
As many projects, \popasm\ code is distributed in many source files
due to the clear advantages of modular design. The objective of this
chapter is to offer the reader an overview of how \popasm\ code is
organized, that is, what modules have been defined, which functionality
each one offers, and how they interact with one another.

Everything starts at main() function in main.cpp file. Its job is to
start \popasm\ up, parse command line options and so on (details in
Chapter \ref{chapmain}). After main() is done with this initialization
it calls the methods for assembling the input files.

Note that \popasm\ source defines an Assembler class (see Chapter
\ref{chapasm}) which carries out the assembling task via its
AssembleFile() method. This method can be seen as the ``root'' of
\popasm\ execution, for it calls the other vital modules of the
assembler.

One of such modules is the parser (described in Chapter \ref{chapparser}),
which calls the lexical analyzer (see Chapter \ref{chaplex}) to read
the keywords, expressions, etc. from the input files. Such lexical
items are then checked to verify whether or not they follow \popasm\ 
syntax. If they do, the parser then converts them into arguments
(such as memory references) and pass them to the instruction (or
directive) found in the current line.

The Instruction class (explained in Chapter \ref{chapinst}) gets such
arguments and performs additional checking (eg. type checking). Unless
an error is found, the instruction is assembled and the resulting encoding
is returned to the parser.

The paragraphs above rawly descrive how the assembly process works.
However many details have been intentionally ommited in order to
make the explanation simple. Such details are fully covered in other
chapters.

\chapter{What main() does\label{chapmain}}
To be written.

\chapter{The assembler classes\label{chapasm}}
One of \popasm\ greatest concerns is compatibility with existing code.
Such code was written according to syntatic rules of other assemblers.
Most of those rules can be gathered together in a single
assembler (like \popasm\ does), but some of them conflict with one
another. For example, in \tasm\ a variable name stands for its contents
while in \nasm\ it is treated as its offset.

Due to such conflicts, command line options have been created to change
\popasm\ configuration. Nevertheless, there are differences among
assemblers which are better tackled by emulating the target assembler
(for example, operator precedence is slightly different between \popasm\ 
and \nasm.).

In order to do that, a base class from which all assemblers to be emulated
derive has been defined. It contains all methods required to perform
the assembly task. The emulated assemblers then inherit from the base
class and overrides the necessary methods to achieve the desired compatibility.
The next sections describe the Assembler base class and PopAsm, the
only supported assembler at the time of this writing.

\section{Assembler base class}
The definition of this class can be found in asmer.h file. It contains four
integer members:

\begin{itemize}
\item InitialMode --- contains the mode (16 or 32 bits) the assembler must
be set to before assembling the source file. In other words, this is the
default operating mode the source will be assembled at, unless explicitly
changed (eg. via a BITS directive);

\item CurrentMode --- Keeps track of which mode the assembler is operating
at. This is the member changed by the BITS directive.

\item CurrentPass --- This member is used to store which pass the assembler
is in. It may be necessary to perform several passes to assemble a single file.

\item CurrentOffset --- stores the memory address the next instruction or
variable will occupy in memory.
\end{itemize}

The methods GetCurrentMode(), SetCurrentMode() and GetCurrentOffset() allow
the interaction with those members.

The only constructor for this class takes the initial mode as its argument.
It initializes the InitialMode member to this value and sets CurrentPass
to zero. Note that CurrentMode and CurrentOffset do not need to be
initialized because the assembler will do that properly at the beginning of
each pass.

The other methods are PerformPass() and AssembleFile(). The latter takes an
input file (InputFile is described in Chapter \ref{chapinp}) as argument and
assembles it. This is done by calling PerformPass() as many times as needed.

PerformPass(), in turn, performs a single pass through the input file and
returns true if another pass is needed or false if assembly success or failure
has already been determined. This method resets the CurrentOffset and CurrentMode
members and parses the lines from the input file, one by one (by means of
Parser::ParseLine() method. See chapter \ref{chapparser} for details). After
parsing a line, this method updates the CurrentOffset accordingly.

\section{PopAsm class}
This is currently the only class that inherits from Assembler, defined in
files popasm.cpp and popasm.h. The constructor takes the initial mode as
its only argument. It also initializes the operator, encloser, instruction
and directive tables. The destructor is responsible for releasing all memory
from the operator and encloser tables (the other tables use static memory and
thus do not need to be destroyed).

All the heavy work is performed in the base class by its default AssemblyFile()
and PerformPass() methods.

\chapter{The Parser\label{chapparser}}
The Parser class can be found in parser.h. The parser is intended to play a
support role for the Assembler, Instruction and Directive classes via its
methods ParseLine() and ParseArguments(). It also checks whether each line
of code conforms to \popasm\ syntax.

Its constructor takes an InputFile object as argument (see Chapter \ref{chapinp}
for details about that class). That is the file the non-static methods act on.
The next sections discuss how the important methods work.

\section{ParseLine() method}
This is the main function of Parser. It is called by Assembler once for each line
of source code. It calls ReadLine() method to read the current line and split
it into a vector of Tokens. These tokens are then identified and the state machine
sketched in Figure \ref{SYNTAXSM} is run (via the \emph{switch} statement).

\begin{figure}[h]
\begin{center}
\begin{picture}(300, 100)(0,0)
\put (  0, 20){\vector(1, 0){25}}

\put ( 40, 17){\scriptsize{INIT}}
\put ( 50, 20){\circle{50}}
\put ( 75, 20){\vector(1, 0){60}}
\put ( 85, 25){\scriptsize{Command}}

\put (150, 17){\scriptsize{CMD}}
\put (160, 20){\circle{50}}
\put (185, 20){\vector(1, 0){60}}
\put (190, 25){\scriptsize{Rest of line}}

\put (260, 17){\scriptsize{END}}
\put (270, 20){\circle{55}}
\put (270, 20){\circle{50}}

\put ( 90, 72){\scriptsize{LABEL}}
\put (105, 75){\circle{50}}
\put ( 67, 37){\vector(1, 1){20}}
\put (122, 56){\vector(1, -1){20}}
\put ( 50, 50){\scriptsize{Label}}
\put (135, 50){\scriptsize{Command}}
\end{picture}
\caption{Syntax state machine}
\label{SYNTAXSM}
\end{center}
\end{figure}

The state machine has four states:

\begin{itemize}
\item INITIAL (INIT) --- is the initial state of the state machine. At
this point nothing has been parsed so far.
\item LABEL --- A label has been read (optionally followed by a colon).
\item COMMAND (CMD) --- The command of this line was found. It may be
either a machine instruction or an assembler directive.
\item END --- Final state.
\end{itemize}

The reader might notice the arguments after the command were not parsed.
The reason for that is that some assembler directives (such as ASSUME)
do not work with usual arguments, but only with the raw sequence of Tokens.
Most commands, however, need the parser to hand them their arguments
after they had been processed. For example, a ``MOV AL,2+2'' instruction
does not ``care'' whether the user typed in just a number or a complex
arithmetic expression; it must be given only the final result.

In order to simplify the implementation of such commands Parser has a
public static method, ParseArguments(), described next.

\section{ParseArguments() method}
This method converts a raw sequence of Tokens into arguments machine
instructions can work with. This conversion cannot be performed within
ParseLine() method because some commands require special parsing.

The required arguments for this method are two iterators into a vector
of Tokens and a vector to output the Arguments to. Those iterators
define an open interval, so that the second iterator points to the
first position past the last Token.

That interval is sequencially searched for a comma or end of interval.
When either is found the first argument Tokens will have been delimited.
The second argument Tokens are the ones from this point on to the next
comma or end of interval, and so on.

Every time the Tokens for a particular argument are determined the
EvaluateExpression() method is called. It performs all arithmetic
operations and returns an Expression (see Chapter \ref{chapexps}).
This Expression is then converted to an Argument and stored in the
vector of Arguments. This conversion is carried out by a static
method of class Argument (details in Chapter \ref{chapargs}),
named MakeArgument().

\section{EvaluateExpression() method}


\section{ReadLine() method}





\chapter{Instructions\label{chapinst}}







\chapter{Lexical Analyzer\label{chaplex}}

\chapter{Definitions\label{chaprefs}}

\chapter{Numbers\label{chapnums}}

\chapter{Input Files\label{chapinp}}

\chapter{Hash Tables\label{chaphash}}

\chapter{Numbers again\label{chapnumsagain}}
This chapter shows how \popasm\ implements its huge numbers classes.
They provide the developer unlimited bit sized numbers to operate
with, if there is enough memory to hold them. Three classes have
been defined to acomplish that: \emph{NaturalNumber}, \emph{IntegerNumber}
and \emph{RealNumber}. They can be found in files \emph{number.h} and
\emph{number.cpp}, and will be discussed in the next sections.

\section{NaturalNumber}
An object of this class can hold \emph{any} non-negative integer number,
no matter how huge it is, with no precision loss. It is implemented
as a vector of \emph{Words} (16-bit unsigned integers). For the rest of
this section, let $V$ be this vector and $n$ be the number it represents.

\subsection{Internal representation\label{intrep}}
The least significant bits of $n$ are stored at the first positions of $V$
(that is, $n$ is encoded in little-endian notation).
For example\footnotemark, if $n = 123456789h$ then $V$ will look as in Table \ref{tabexv}.

\begin{table}[h]
\begin{center}
\begin{tabular}[h]{|c|c|c|}
V[2]  & V[1]  & V[0]\\
\hline
0001h & 2345h & 6789h\\
\hline
\end{tabular}
\label{tabexv}
\caption{Vector $V$ holding $n = 123456789h$}
\end{center}
\end{table}
\footnotetext{Note that $V$ was displayed in reverse order. This
convention will be used in this section to make numbers and math
operations easier to read.}
Another important property of $V$ is that it never contains trailing
zeroes (that is, its last element -- the most significant \emph{Word}
-- is never zero). Obviously, the first \emph{Words} may be zero,
provided they are followed by a non-zero \emph{Word}. Hence Table \ref{tabwrong}
shows an invalid representation of $n$ (due to a trailing zero at
$V[3]$).

\begin{table}[h]
\begin{center}
\begin{tabular}[h]{|c|c|c|c|}
V[3]  & V[2]  & V[1]  & V[0]\\
\hline
0000h & 0001h & 2345h & 6789h\\
\hline
\end{tabular}
\label{tabwrong}
\caption{Invalid representation of $n = 123456789h$}
\end{center}
\end{table}

Finally, $n$ is zero if and only if $V$ is empty.

\subsection{The constructors}
\emph{NaturalNumber} has two constructors. One of them takes a
\emph{Dword} (32-bit unsigned integer) that defaults to be zero,
splits it in two \emph{Words} and place them into $V$ (but they
are checked to prevent trailing zeroes, as mentioned in subsection
\ref{intrep}).

The other constructor reads a \emph{NaturalNumber} from a string. It
first calls \emph{SimplifyString()} to detect the sign\footnote{
The reader might argue that a \emph{NaturalNumber} can't be
negative and thus it would not make sense to detect the sign of the
number, but recall that \emph{SimplifyString()} is not a method of
\emph{NaturalNumber}; it is also used by the other classes (such as
\emph{IntegerNumber}) which can hold negative numbers as well.} and
base the number was written in and to strip trailing underscores.

The number is then calculated by the usual algorithm.

\begin{quotation}
\noindent PartialNumber $\leftarrow$ 0;\\
\\
while there are remaining digits do\\
\indent NextDigit $\leftarrow$ ReadNextDigit();\\
\indent PartialNumber $\leftarrow$ PartialNumber * BASE;\\
\indent PartialNumber $\leftarrow$ PartialNumber + NextDigit;\\
end while
\end{quotation}

The other argument, \emph{ForcedBase}, tells the constructor to use
the specified base instead of the default one. It is used by one of
the constructors of RealNumber class (see section \ref{secrealnum}).

\subsection{Performing math operations}
\subsubsection{Addition}
\emph{NaturalNumber} class performs addition by adding \emph{Words}
of both numbers, two at a time, and propagating the carry forward. For
example, table \ref{tabadd} shows how we could add $12345678h$ and
$32055h$.

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|c}
V[1]  & V[0]  &\\
\hline
1234h & 5678h & A\\
\hline
0003h & 2055h & B\\
\hline
1237h & 76CDh & A + B\\
\hline
\end{tabular}
\label{tabadd}
\caption{Simple example of addition}
\end{center}
\end{table}

Table \ref{tabadd} shows a rather simple example, because both numbers
have the same size in \emph{Words} and there is no carry. Let's consider
another example where both of these issues occur (table \ref{tabadd2}).

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|c|c}
V[2]  & V[1]  & V[0] & \\
\hline
      & FFFFh & 9314h & A\\
\hline
      &       & 8000h & B\\
\hline
0001h & 0000h & 1314h & A + B\\
\hline
\end{tabular}
\label{tabadd2}
\caption{A more complex example of addition}
\end{center}
\end{table}

As table \ref{tabadd2} shows, the first addition of \emph{Words} at V[0]
generates a carry, because $9314h + 8000h = 11314h$, which is beyond the
$FFFFh$ limit for a \emph{Word}. Thus the V[0] of the result will hold the
16 least significant bits of $11314h$ (that is, $1314h$) and a carry will be
propagated to the next stage of the addition.

The next stage would be to add $FFFFh$ and an empty vector position
(which can be interpreted as zero) and the carry due to the previous stage.
The result will be $FFFFh + 0000h + 1 = 10000h$. This value, too, is greater
than $FFFFh$, so its 16 least significant bits ($0000h$) are stored at V[1] and
a new carry is generated, but we are already done with adding $A$ and $B$, so
the carry is appended to $A$, yielding $A + B$ as shown in table \ref{tabadd2}.

Note that some cells of table \ref{tabadd2} are blank, not zero. That's because
$A$ has a two \emph{Words} long vector, while $B$ has a single-element vector of
\emph{Words} to represent its value. We could not add trailing zeroes to either
number, otherwise we would be violating the restriction that states that no
number may have trailing zeroes at its vector (see section \ref{intrep}).

From the above discution, it is easy to get to the algorithm \popasm\ uses to
add two \emph{NaturalNumbers} (the sum is placed onto $A$, that is, this is
the algorithm for \emph{NaturalNumber::operator+=}):

\begin{quotation}
\noindent 01 Dword temp $\leftarrow$ 0;\\
02 \noindent \\
03 \noindent if A.SizeInWords() $<$ B.SizeInWords()\\
04 \indent   match $A$ and $B$ sizes by pushing back zeroes\\
05 \noindent end if\\
06 \noindent \\
07 \noindent for i $\leftarrow$ 0 to A.SizeInWords() - 1 do\\
08 \indent   temp $\leftarrow$ Dword(A[i]) + Dword(B[i]) + temp;\\
09 \indent   A[i] $\leftarrow$ temp[0..15];\\
10 \indent   temp $\leftarrow$ temp[16..31];\\
11 \noindent end for
\end{quotation}


\section{IntegerNumber}
\section{RealNumber\label{secrealnum}}



\chapter{Operators\label{chapops}}

\chapter{Registers\label{chapregs}}

\chapter{Expressions\label{chapexps}}

\chapter{Arguments\label{chapargs}}


\appendix

\chapter{Contacting info\label{CONTACT}}
Up to date info about contacting \popasm\ team can be found at \popasm\
homepage, at http://popasm.sourceforge.net.

\begin{thebibliography}{12}
\bibitem{UFES} Federal University of Esp\'{\i}rito Santo -- ES -- Brazil.
http://www.ufes.br
\bibitem{PUC} Pontificial Catholic University of Rio de Janeiro -- RJ
-- Brazil. http://www.puc-rio.br
\bibitem{SF} SourceForge. http://sorceforge.net
\bibitem{ADV}Advogato. http://www.advogato.com
\bibitem{ADVPRSN}H\'elcio Mello's personal page at Advogato.
http://www.advogato.com/person/helcio
\bibitem{FRESH}FreshMeat. http://www.freahmeat.net
\bibitem{FREITAS}Professor PhD S\'ergio A. A. Freitas.
http://www.inf.ufes.br/\~{}sergio.
\bibitem{AUTOCONF}GNU Autoconf.
http://www.gnu.org/software/autoconf/autoconf.html
\bibitem{AUTOMAKE}GNU Automake.
http://www.gnu.org/software/automake/automake.html
\bibitem{MAKE}GNU Make. http://www.gnu.org/software/make/make.html
\bibitem{FPUMAN}FPU Programming on x86 FPU's
\end{thebibliography}

\end{document}
